On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification

### 2021-03-29 paper/cav_determinisim_main.tex
Added CHANCE to headnote on the IEEE paper's version
modified method diagram

Digital Object Identifier ....
}}}
% \affil{Trustworthy Systems Laboratory, University of Bristol, Bristol, UK}
% \affil{University of West of England, Bristol, UK}
% \affil{Bristol Robotics Laboratory, Bristol, UK}
%
\markboth{IEEE TRANSACTIONS ON INTELLIGENT TRANSPORTATION SYSTEMS VOL. ..., NO. ..., date...}{Ghobrial \MakeLowercase{\textit{et al.}}: On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification}
%
\maketitle
%
\begin{abstract}
\noindent 
Game engines are increasingly used as simulation platforms by the autonomous vehicle (AV) community to develop vehicle control systems and test environments. 
%
A key requirement for simulation-based development and verification is determinism, since a deterministic process will always produce the same output given the same initial conditions and event history. 
%
Thus, in a deterministic simulation environment, tests are rendered repeatable and yield simulation results that are trustworthy and straightforward to debug. 
%
However, game engines are seldom deterministic. 
%
This paper reviews and identifies the potential causes of non-deterministic behaviours in game engines. 
%
A case study using CARLA, an open-source autonomous driving simulation environment powered by Unreal Engine, is presented to highlight its inherent shortcomings in providing sufficient precision in experimental results. 
%
Different configurations and utilisations of the software and hardware are explored to determine an operational domain where the simulation precision is sufficiently low i.e.\ variance between repeated executions becomes negligible for development and testing work. 
%
Finally, a method of a general nature is proposed, that can be used to find the domains of permissible variance in game engine simulations for any given system configuration.

@@

% This was calculated by averaging the variance between all of the different runs through out the simulation time for all of the agents, then taking the square root to obtain the total mean devaiton. Note the dotted red line in the plot shows the millimeter (mm) level. Below this level we assume that the simulation variance is low enough that it would not affect our assertion checking for V\&V of AV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/Methodology_Diagram_v7.pdf}
    \caption{Stages of the method proposed to determine the variance of a simulation.}
    \label{method_diagram}
\end{figure*}

>>>>

Digital Object Identifier ....
}}}
% \affil{Trustworthy Systems Laboratory, University of Bristol, Bristol, UK}
% \affil{University of West of England, Bristol, UK}
% \affil{Bristol Robotics Laboratory, Bristol, UK}
%
\markboth{IEEE TRANSACTIONS ON INTELLIGENT TRANSPORTATION SYSTEMS VOL. ..., NO. ..., date...}{Chance, Ghobrial \MakeLowercase{\textit{et al.}}: On Determinism of Game Engines used for Simulation-based Autonomous Vehicle Verification}
%
\maketitle
%
\begin{abstract}
\noindent 
Game engines are increasingly used as simulation platforms by the autonomous vehicle (AV) community to develop vehicle control systems and test environments. 
%
A key requirement for simulation-based development and verification is determinism, since a deterministic process will always produce the same output given the same initial conditions and event history. 
%
Thus, in a deterministic simulation environment, tests are rendered repeatable and yield simulation results that are trustworthy and straightforward to debug. 
%
However, game engines are seldom deterministic. 
%
This paper reviews and identifies the potential causes of non-deterministic behaviours in game engines. 
%
A case study using CARLA, an open-source autonomous driving simulation environment powered by Unreal Engine, is presented to highlight its inherent shortcomings in providing sufficient precision in experimental results. 
%
Different configurations and utilisations of the software and hardware are explored to determine an operational domain where the simulation precision is sufficiently low i.e.\ variance between repeated executions becomes negligible for development and testing work. 
%
Finally, a method of a general nature is proposed, that can be used to find the domains of permissible variance in game engine simulations for any given system configuration.

@@

% This was calculated by averaging the variance between all of the different runs through out the simulation time for all of the agents, then taking the square root to obtain the total mean devaiton. Note the dotted red line in the plot shows the millimeter (mm) level. Below this level we assume that the simulation variance is low enough that it would not affect our assertion checking for V\&V of AV simulations purposes, that it is sufficient to assume the tests are deterministic and repeatable. This line can be moved up or down and it is entirely up to verification engineers where they want to define the level below which tests' variance is negligibale and it is suffcient to assume that tests are deterministic.
%
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{Other/Figures/Methodology_Diagram_v8.pdf}
    \caption{Stages of the method proposed to determine the variance of a simulation.}
    \label{method_diagram}
\end{figure*}
